



```{r}
knitr::opts_chunk$set(fig.width=6, fig.height=6) 
```

# Generate species distributions


```{r}
source("setParams.R")
```

Set size of area over which to simulate species. Default is for a square of 300 by 300


```{r}
dim <- c(300,300)
```

Decide on an average number of individuals per cell. The default is 0.05


```{r}
indivs <- 0.05
lambda <- log(indivs)
```

Decide on the strength of environmental gradient. Default is 1.2


```{r}
env.beta <- 1.2
```

The expected number of individuals across the whole area can be roughly estimated as:


```{r}
EN <- exp(lambda + (env.beta*0.5)) * (dim[1]*dim[2])
print(EN)
```

Choose how much autocorrelation there is in the species distribution. Default is 0.05, increasing simulates decreased spatial autocorrelation, decreasing this value gives increased autocorrelation


```{r}
kappa <- 0.05
```

Run the data generation script and view the outputs


```{r}
source("genData.R")

dat1 <- suppressMessages(genData(dim = dim, lambda = lambda, env.beta = env.beta, plotdat = TRUE, seed = NULL, sigma2x = sigma2x, kappa = kappa))
##add points to show individuals
points(dat1$xy, pch = 20, cex = 0.1, col = "black")
```

Calculate the number of individuals simulated


```{r}
print(nrow(dat1$xy))
```

# Generate a citizen science survey

Decide on the maximum probability of a citizen scientist observing an individual. Default is 0.5


```{r}
maxprob <- 0.5
```

By default the citizen science survey is biased with one region of the area having a higher probability of sampling
(up to the maxprob value you just specified).

You can decide, should the pattern in citizen science bias be correlated with the environmental gradient? Default is FALSE


```{r}
correlated <- FALSE
```


```{r}
##Set up bias pattern

source("Generate strata levels Lam.R")
strata1 <- genStrataLam(dat1$Lam, strata = strata, rows = rows, cols = cols, plot = FALSE)
source("addSpatialBias.R")
biasfield <- addSpatialBias(strata1, maxprob = probs[1], correlated = correlated, rho = rho)
biascov <- xtabs(covariate ~ y + x, data = biasfield)
```


```{R}
##Generate thinned citizen science data

source("thinData.R")
thin1 <- thinData(dat1, biasfield)
thin1$env <- dat1$gridcov[as.matrix(thin1[,2:1])]
thin1$bias <- biascov[as.matrix(thin1[,2:1])]

unstructured_data <- thin1
```

Visualise citizen science data


```{r}
image.plot(list(x=dat1$Lam$xcol, y=dat1$Lam$yrow, z=t(dat1$rf.s)), main='Thinned unstructured data', asp=1, cex.main =1) 
points(unstructured_data$x, unstructured_data$y, pch = 20)
```

Calculate number of citizen science observations:


```{r}
n_citsci <- nrow(unstructured_data)
print(n_citsci)
```

## Generate a planned survey

Planned surveys are assumed to be spatially unbiased.

How many survey locations should there be? Default is 150


```{r}
nsamp <- 150
```

How large should surveys be (i.e. how many grid cells?). Default is 1


```{r}
qsize <- 1
```


```{r}
source("sampleStructured.R")

structured_data <- sampleStructured(dat1, biasfield, nsamp = nsamp, plotdat = FALSE, qsize = qsize)
```

Visualise structured data


```{r}
image.plot(list(x=dat1$Lam$xcol, y=dat1$Lam$yrow, z=t(dat1$rf.s)), main='Structured data', asp=1, cex.main = 1) 
points(structured_data$x,structured_data$y, pch = 21, bg = structured_data$presence, col = "black")
par(xpd = TRUE)
legend(-10,360,c("Absence  ", "Presence  "), pch = 21, col = "black", pt.bg = c(0,1), cex = 0.8)
```

# Run model for citizen science data

Decide whether the model should include a covariate for bias (i.e. the modeller knows what is causing bias in citizen science observations). Default is TRUE


```{r}
biasknown <- TRUE
```


```{r}
if(biasknown == TRUE){
    source("Run models unstructured bias covariate.R")
    mod_citsci <- suppressWarnings(unstructured_model_cov(unstructured_data, dat1, biasfield, dim = dim, plotting = FALSE, 
                                                          biascov = biascov))
} else {
    source("Run models.R")
    mod_citsci <- suppressWarnings(unstructured_model(unstructured_data, dat1, biasfield, dim = dim, plotting = FALSE))
    }
```


```{r}
source("validation_function.R")
validation_citsci <- validation_function(result=mod_citsci[[2]], resolution=c(10,10), join.stack=mod_citsci[[1]], 
                                         model_type="unstructured", unstructured_data = unstructured_data, dat1 = dat1, 
                                         summary_results=T, qsize = 1, absolute=FALSE, dim = dim, plotting = TRUE)

```

# Run model for planned survey data


```{r}
source("Run models structured.R")
mod_planned <- suppressWarnings(structured_model(structured_data, dat1, biasfield, plotting = FALSE))
```


```{r}
source("validation_function.R")
validation_planned <- validation_function(result=mod_planned[[2]], resolution=c(10,10), join.stack=mod_planned[[1]], 
                                          model_type="structured", structured_data = structured_data, dat1 = dat1, 
                                          summary_results=T, qsize = 1, absolute=FALSE, dim = dim, plotting = TRUE)

```

# Run integrated model


```{r}
if(biasknown == TRUE){
    source("Run models joint covariate for bias.R")
    mod_integrated <- suppressWarnings(joint_model_cov(structured_data, unstructured_data, dat1, biasfield, 
                                                       resolution = c(10,10), biascov = biascov))
} else {
    source("Run models joint.R")
    mod_integrated <- suppressWarnings(joint_model(structured_data, unstructured_data, dat1, biasfield))
}
```


```{r}
source("validation_function.R")
validation_integrated <- validation_function(result=mod_integrated[[2]], resolution=c(10,10), join.stack=mod_integrated[[1]], 
                                             model_type="joint", unstructured_data = unstructured_data, 
                                             structured_data = structured_data,dat1 = dat1, summary_results=T, absolute = FALSE,
                                             dim = dim, plotting = TRUE)

```

## Model comparison table


```{r}
modeldf <- data.frame(model_type = c("Citizen science", "Planned", "Integrated"), MAE = c(as.numeric(validation_citsci$`Proto-table`)[2],as.numeric(validation_planned$`Proto-table`)[2],as.numeric(validation_integrated$`Proto-table`)[2]),
                      Correlation = c(validation_citsci$correlation,validation_planned$correlation,validation_integrated$correlation),
                     Env_coeff = c(c(mod_citsci$result$summary.fixed[2,1],mod_planned$result$summary.fixed[2,1],mod_integrated$result$summary.fixed[3,1])))
modeldf
```
